Towards Detecting Stalkerware Through Network Traces

### 2019-03-19 paper/main.tex
Consolidating app families section

\section{The Stalkerware Ecosystem}
Some spyware has a notification that the app is running, but that can be 
disabled. Those ones don't hide their icon, but you might have to go to All 
Apps to find it.
\subsection{The Practice of ``Reskinning''}
\label{reskinning}
As previously mentioned, spyware has a high turnover rate on the Google Play 
store, which leads to a practice we refer to as “reskinning:” a single app will 
be posted under many names, icons, and manufacturer names. Presumably, 
manufacturers do this because spyware gets reported frequently and taken down - 
we have already seen one instance of a spyware app getting removed from the 
Play store while we were examining it. Overt stalkerware which is not available 
through Google Play has also turned out to use reskinning, for reasons we 
haven’t yet discovered. Apps can frequently be identified as copies of each 
other by observing their dashboards (where users view the communications and 
locations of the target phones), download sites, and advertisements. Several 
groups of off-store apps have nearly identical websites, in terms of 
organization, color, fonts, placement of buttons, prices, and other factors. 
For example, the “ILFMobile group” consists of three different companies 
(ILFMobile, PowerLine Group, and PhoneSpector) that claim to make a total of 
four different apps. Further investigation showed that all three companies had 
the same physical address in addition to identical websites. Unfortunately, 
they also had identical bugs in their payment methods, so we were unable to 
purchase the apps and verify that their network traces were identical. We were 
able to check the network signatures for the rest of the app groups in the list 
below (with the exception of one app which we are still confirming). These 
reskinned programs are potentially a boon to our investigation as the identical 
network signatures make it more likely that one signature generalizes across 
multiple apps.
We have identified the following groups of applications that appear to be 
reskinned versions of each other.
\begin{enumerate}
	\item \textbf{SpyToMobile group}: Employee Work Spy by Piter Cline, SMS 
	Tracker by TheHar, Mobile Tracking by Antwat, Cell Phone Tracker by Susdel, 
	and Texting, Chat, 
	Phone Spy by MaxLo. All of these apps download the same program, which is 
	called SpyToMobile. Any app from this group installed after the first 
	states ``SpyToMobile is already installed.''
	\item \textbf{ILFMobile group}: HighsterMobile, PhoneSpector, Auto Forward 
	Spy, Surepoint Spy, and EasySpy. We were unable to purchase these and 
	ensure that they have the same signatures due to a bug in their payment 
	system that prevented us from purchasing them.
	\item \textbf{Codero group}: HelloSpy and TheTruthSpy have identical 
	network signatures as well as nearly identical websites.
	Spyzie group: Spyzie and SpyMyFone also have identical network signatures 
	as well as identical websites.
	\item \textbf{CallSMSTracker group}: Call and Message Tracker by Remote 
	Karath and Cell Tracker by Trackme. These apps have identical dashboards 
	and contact the same domains.
\end{enumerate}
We cannot guarantee that we have identified all of the members of any 
particular group. If we have not, it will make other members easier to 
identify, since signatures we have already found will scale to identify more 
apps than we had time to download and manually examine.

@@

App manufacturers appear to be well aware of their competition. When the term 
``hello spy'' is typed into Google Search, ads appear labeled ``Hellospy | See 
All Phone Data Remotely | mSpy.com'' and ``HelloSpy - Free Download | \#1 Phone 
Tracker Software\&App | spymyfone.com.'' Clicking on these leads the user to 
the download pages for mSpy and SpyMyFone respectively, not the site for 
HelloSpy. Based on network traces and website appearances, we do not believe 
any of these three apps are made by the same manufacturer. Spyware groups also 
create “review” websites that review their own products, posing as third party 
spyware enthusiasts. For example, a member of the ILFMobile group reviews the 
“top 5 cell phone spy apps,” all of which are ILFMobile’s own apps, on a 
website called “bestcellphonespyapps.com.” We haven’t determined what advantage 
spyware manufacturers derive from appearing to make more apps and competing 
with themselves.

@@

However, some characteristics of the stalkerware apps themselves may aid us in 
detecting them. For example, similar to malware, some groups of apps appear to 
be repackaged versions of each other. For example, PhoneSpector, 
HighsterMobile, AutoForward, EasySpy and Surepoint Spy all have separate 
websites, but the appearance of each of these sites is nearly identical. Upon 
closer inspection, although multiple companies are listed as the makers of the 
apps (Powerline Group Corp., ILF Mobile Apps, and PhoneSpector LLC), all have 
the same physical address listed. At least one blog that is supposedly run by 
an independent party reviewed apps that coincidentally were all made by the 
same company\cite{noauthor_best_nodate}. Other app sites appear to be similar 
to each other in other ways, 
with many having a large panel with a darkened image background extolling the 
features of the app above two buttons, one for a "live demo" and the other to 
"buy now." This is somewhat circumstantial evidence - it is entirely possible 
that these manufacturers simply chose similar web templates, or copied each 
other's site in an attempt to undercut competition. We therefore confirmed our 
suspicions by checking if the traffic signatures were also similar. In multiple 
cases, this was true (see section \ref{app_families}). App identification may 
be easier to scale since multiple groups of stalkerware apps are in fact made 
by only a few manufacturers.

@@

\subsection{App ``Families''}
\label{app_families}
After our investigation on these apps, we can confirm that, at least from the 
few that we have selected, many stalkerware apps are related in some fashion to 
each other. The most notable being the group of apps stated above, and a single 
app mentioned before called SpyToMobile. SpyToMobile is an interesting case 
because it fits the description of a generic stalkerware app so closely that it 
is the perfect example of spyware on the Play Store. Moreover, this single app 
has been repackaged multiple times. We have discovered that there are at least 
6 copies of the app on the Play store, all with different developers and 
different names. While this is still circumstantial evidence that most 
stalkerware often contains copies of each other, we can at least conclude that 
our forty-five apps constitute a larger percentage than anticipated.

PUT NOTE HERE ABOUT THE FAMILIES THAT WE DID NOT CONFIRM VIA TRAFFIC, BUT 
INSTEAD VIA LOOKING AT THE WEBSITES!! Otherwise we're fabricating results.

>>>>

\section{The Stalkerware Ecosystem}
Some spyware has a notification that the app is running, but that can be 
disabled. Those ones don't hide their icon, but you might have to go to All 
Apps to find it.
\subsection{The Practice of ``Reskinning''}
\label{reskinning}
As previously mentioned, spyware has a high turnover rate on the Google Play 
store, which leads to a practice we refer to as “reskinning:” a single app will 
be posted under many names, icons, and manufacturer names. Presumably, 
manufacturers do this because spyware gets reported frequently and taken down - 
we have already seen one instance of a spyware app getting removed from the 
Play store while we were examining it. Overt stalkerware which is not available 
through Google Play has also turned out to use reskinning, for reasons we 
haven’t yet discovered. Apps can frequently be identified as copies of each 
other by observing their dashboards (where users view the communications and 
locations of the target phones), download sites, and advertisements. Several 
groups of off-store apps have nearly identical websites, in terms of 
organization, color, fonts, placement of buttons, prices, and other factors. 
For example, the “ILFMobile group” consists of three different companies 
(ILFMobile, PowerLine Group, and PhoneSpector) that claim to make a total of 
four different apps. Further investigation showed that all three companies had 
the same physical address in addition to identical websites. Unfortunately, 
they also had identical bugs in their payment methods, so we were unable to 
purchase the apps and verify that their network traces were identical. Other 
app sites appear to be similar to each other in other ways, 
with many having a large panel with a darkened image background extolling the 
features of the app above two buttons, one for a "live demo" and the other to 
"buy now." This is somewhat circumstantial evidence - it is entirely possible 
that these manufacturers simply chose similar web templates, or copied each 
other's site in an attempt to undercut competition. We therefore confirmed our 
suspicions by checking if the traffic signatures were also similar. These 
reskinned programs are potentially a boon to 
our investigation as the identical network signatures make it more likely that 
one signature generalizes across multiple apps. We have identified the 
following groups of applications that appear to be reskinned versions of each 
other.
\begin{enumerate}
	\item \textbf{SpyToMobile group}: Employee Work Spy by Piter Cline, SMS 
	Tracker by TheHar, Mobile Tracking by Antwat, Cell Phone Tracker by Susdel, 
	and Texting, Chat, 
	Phone Spy by MaxLo. All of these apps download the same program, which is 
	called SpyToMobile. Any app from this group installed after the first 
	states ``SpyToMobile is already installed.''
	\item \textbf{ILFMobile group}: HighsterMobile, PhoneSpector, Auto Forward 
	Spy, Surepoint Spy, and EasySpy. We were unable to purchase these and 
	ensure that they have the same signatures due to a bug in their payment 
	system that prevented us from purchasing them.
	\item \textbf{Codero group}: HelloSpy and TheTruthSpy have identical 
	network signatures as well as nearly identical websites.
	Spyzie group: Spyzie and SpyMyFone also have identical network signatures 
	as well as identical websites.
	\item \textbf{CallSMSTracker group}: Call and Message Tracker by Remote 
	Karath and Cell Tracker by Trackme. These apps have identical dashboards 
	and contact the same domains.
\end{enumerate}
We cannot guarantee that we have identified all of the members of any 
particular group. If we have not, it will make other members easier to 
identify, since signatures we have already found will scale to identify more 
apps than we had time to download and manually examine.

\subsubsection{SpyToMobile}
SpyToMobile is an interesting case 
because it fits the description of a generic stalkerware app so closely that it 
is the perfect example of spyware on the Play Store. Moreover, this single app 
has been repackaged multiple times. We have discovered that there are at least 
6 copies of the app on the Play store, all with different developers and 
different names. While this is still circumstantial evidence that most 
stalkerware often contains copies of each other, we can at least conclude that 
our forty-five apps constitute a larger percentage than anticipated.

@@

App manufacturers appear to be well aware of their competition. When the term 
``hello spy'' is typed into Google Search, ads appear labeled ``Hellospy | See 
All Phone Data Remotely | mSpy.com'' and ``HelloSpy - Free Download | \#1 Phone 
Tracker Software\&App | spymyfone.com.'' Clicking on these leads the user to 
the download pages for mSpy and SpyMyFone respectively, not the site for 
HelloSpy. Based on network traces and website appearances, we do not believe 
any of these three apps are made by the same manufacturer. Spyware groups also 
create “review” websites that review their own products, posing as third party 
spyware enthusiasts. At least one blog that is supposedly run by 
an independent party reviewed apps that coincidentally were all made by the 
ILFMobile group\cite{noauthor_best_nodate}. This blog reviews the 
“top 5 cell phone spy apps,” all of which are ILFMobile’s own apps, on a 
website called “bestcellphonespyapps.com.” We haven’t determined what advantage 
spyware manufacturers derive from appearing to make more apps and competing 
with themselves.

@@

@@
